---
title: "02 chart EU migrant"
author: "Duc-Quang Nguyen"
date: "19 September 2016"
output: html_document
---

## Resources

* [Libre circulation](http://www.touteleurope.eu/les-politiques-europeennes/marche-interieur/synthese/la-libre-circulation-des-personnes.html)
* [SEM libre circulation](https://www.sem.admin.ch/sem/fr/home/themen/fza_schweiz-eu-efta.html) AELE + EU
* To check the graphic!! [Eurostat Non-national population by group of citizenship](http://ec.europa.eu/eurostat/statistics-explained/index.php/File:Non-national_population_by_group_of_citizenship,_1_January_2015_(¹)_YB16.png)

## Related

* [Metrocosm Which EU Country Has the Most Citizens Living Abroad?](http://metrocosm.com/eu-diaspora-map/)
* [Metrocosm The Ins and Outs of Immigration in the European Union](http://metrocosm.com/mapping-eu-immigration/)
* [Intra-european living elsewhere](http://www.economist.com/news/europe/21704813-eus-cherished-free-movement-rights-are-less-secure-they-seem-europes-scapegoat)

## Text snippets

Nationalité des ressortissants. Pour les pays pour lequels procesus naturalisation facile, part d'étrangers. 

* The number of people residing in an EU Member State with citizenship of a non-member country on 1 January 2015 was 19.8 million, representing 3.9 % of the EU-28 population. In addition, there were 15.3 million persons living in one of the EU Member States on 1 January 2015 with the citizenship of another EU Member State. [Eurostat Migration and migrant population statistics](http://ec.europa.eu/eurostat/statistics-explained/index.php/Migration_and_migrant_population_statistics) 

* [Libre circulation et l'Union europénne](http://www.vie-publique.fr/chronologie/chronos-thematiques/libre-circulation-personnes-union-europeenne.html)  En 1985, cinq pays (l’Allemagne, la Belgique, la France, le Luxembourg et les Pays-Bas) décident d’étendre aux ressortissants de pays tiers le bénéfice de la libre circulation sur leur territoire et de créer entre eux un territoire sans frontières, l’Espace Schengen, du nom de la ville luxembourgeoise où est signé le premier accord. En 2015, l’espace Schengen regroupe 26 pays et concerne près de 420 millions d’habitants. 22 pays membres de l’Union européenne ont adhéré aux Accords de Schengen (le Royaume-Uni, l’Irlande, la Bulgarie, la Roumanie, Chypre et la Croatie n’en font pas partie). L’Espace Schengen comprend aussi quatre États non-membres de l’UE (Norvège, Islande, Suisse et Liechtenstein)

* [OECD migration outlook 2016](http://www.oecd-ilibrary.org/social-issues-migration-health/international-migration-outlook-2016_migr_outlook-2016-en)

# Text



```{r setup, include=FALSE}
printTest <- F
translation.file <- 'input/European migration waffe - Sheet1.csv'

iso2.sub <- c('CH', 'DE', 'IT', 'GB', 'ES', 'PT','AT', 'RO')
infile <- "input/data_pop_2016.csv"

library(readr)
library(tidyr)
library(dplyr)
library(magrittr)
library(countrycode)
library(ggplot2)
library(scales)
library(swiMap)
library(swiTheme)
library(forcats)
library(RColorBrewer)
### Getting data in packages
library(eurostat)
library(waffle)
library(htmltools)
library(swiRcharts)

```

```{r load data, echo = F, message = F}
data <- read_csv(infile)

# hack for UK and GR iso code 
data$citizen_code<- gsub("UK", "GB", data$citizen_code)
data$geo_code<- gsub("UK", "GB", data$geo_code)
data$citizen_code<- gsub("EL", "GR", data$citizen_code)
data$geo_code<- gsub("EL", "GR", data$geo_code)

eu.ct <- c(eu_countries[,1], efta_countries[,1])
eu.ct <- gsub("UK", "GB", eu.ct)
eu.ct <- gsub("EL", "GR", eu.ct)

# translation
txt <- read.csv(translation.file, row.names = 1, stringsAsFactors = F)
# discard incomplete translations
cidx <- unique(which(txt =="" | is.na(txt), T)[,2])
if(length(cidx > 0)) {
  warning(paste(colnames(txt)[cidx], collapse = "\t"), " languages will be discarded!", "\n")
  txt <- txt[,-cidx, drop = F]
}
colnames(txt)

# Some test
if(printTest) {
data %>% filter(geo_code == "CH") %>% as.data.frame()
data %>% filter(geo_code %in% iso2.sub) %>% group_by(citizen_code, citizen) %>%
  summarise(tot = sum(values)) %>% ungroup() %>% arrange(desc(tot)) %>% head(20)  
}

## EU stats

keyn <- data %>% filter(geo_code %in% eu.ct)


```
### Compute the sum for EU-28 + EFTA

```{r get EU & EFTA totals, results='asis'}

## for all EU & EFTA countries check how complete is the citizen data
eu.test <- data %>% filter(geo_code %in% eu.ct) %>% group_by(geo_code) %>%
  summarise(forNat = length(values)) %>% ungroup() %>% as.data.frame()

cat("### countries missing !!!!")
cat(eu.ct[!eu.ct %in% eu.test$geo_code])

# create a special structure to plot as waffle with the sums of all EU & EFTA countries
eu <- data %>% filter(geo_code %in% eu.ct) %>% group_by(citizen, citizen_code) %>%
  summarise(values = sum(values)) %>% ungroup()

eu.sums <- data %>% filter(geo_code %in% eu.ct) %>% group_by(geo) %>% 
  summarise(
    TOTAL = first(TOTAL),
    NAT = first(NAT),
    FOR = first(FOR),
    EFTA_FOR = first(EFTA_FOR),
    EU28_FOR = first(EU28_FOR)
  ) %>% ungroup()

eu.2<- t(as.data.frame(colSums(eu.sums[,-1])))
rownames(eu.2) <- NULL
eu <- cbind(eu, eu.2)

eu$geo <- "EU_EFTA"
eu$geo_code <- 'EU'

eu$EUEFTA <- eu$EFTA_FOR + eu$EU28_FOR
eu %<>% select(-EFTA_FOR, -EU28_FOR) %>% arrange(desc(values))



eut <- data %>% filter(geo_code %in% eu.ct) %>% group_by(citizen) %>%
  summarise(totNat = sum(values)) %>% ungroup() %>% arrange(desc(totNat)) %>%
  mutate(TOT = sum(totNat))

eut$citizen_code <- as.character(unlist(data[match(eut$citizen, data$citizen), 'citizen_code']))
eut$share <- eut$totNat / eut$TOT
eut$euefta <- eut$citizen_code %in% eu.ct
eut$continent <- countrycode(eut$citizen_code, "iso2c", "continent")

# hack
eut[which(eut$citizen_code %in% c("XK", "TW")), 'continent'] <- c("Europe", "Asia")
if(any(is.na(eut$continent))) {stop()}


if(printTest) {
  head(eut, 20)
}

# library(treemap)
# library(highcharter)
# 
# dff <- treemap(eut, draw = T, index = c("continent", "citizen"), vSize = "totNat", type = "index")
```



```{r prepare data for waffle chart, echo = F}

# compute the sum EFTA + EU-28 foreigners by country
data$EUEFTA <- data$EFTA_FOR + data$EU28_FOR
data %<>% select(-EFTA_FOR, -EU28_FOR)

#dd <- data %>% filter(geo_code == "IT") %>% arrange(desc(values))

mergeCit <- function(df, n = 6, base = 1000) {
  df %<>% arrange(desc(values))
  topN <- df %>% select(citizen_code) %>% head(n) %>% unlist(use.names = F)
  
  # get the total for EUEFTA and non-EU/EFTA countries
  euefta <- df %>% filter(citizen_code %in% topN, citizen_code %in% eu.ct) %>% 
    summarise(tot = sum(values)) %>% unlist()
  non_euefta <- df %>% filter(citizen_code %in% topN, !citizen_code %in% eu.ct) %>% 
    summarise(tot = sum(values)) %>% unlist()

  ddd <- df %>% select(citizen_code, geo_code, values, TOTAL, FOR) %>% filter(citizen_code %in% topN)
  
  ddd <- rbind(ddd, 
        data.frame(citizen_code = "OU", geo_code = ddd$geo_code[1], 
                   TOTAL = ddd$TOTAL[1], FOR = ddd$FOR[1], values = df$EUEFTA[1] - euefta),
        data.frame(citizen_code = "OT", geo_code = ddd$geo_code[1],  
                   TOTAL = ddd$TOTAL[1], FOR = ddd$FOR[1], values = df$FOR[1] - non_euefta- df$EUEFTA[1])
  )
  
  ddd$share <- round((ddd$values / ddd$TOTAL) * base)
  ddd
}

### Subset data by iso2.sub and merge foreigner citizenships! 
dd <- filter(data, geo_code %in% iso2.sub)
df <- do.call(rbind, by(dd, dd$geo_code, mergeCit))
df <- rbind(df, mergeCit(eu))

# Get all foreign nationality
cit <- df %>% group_by(citizen_code) %>% summarise(tot = sum(share)) %>%
  ungroup() %>% arrange(desc(tot))

colors <- c( "#dbd6db", "#e5dbcd",
            "#ab3d3f", "#666699", "#669999", "#663333", "#336666", "#366096", "#333366", "#996699", "#ac673e", "#3a9736", "#999966", "#666633", "#89a23a", "#368596", "#996666", 
            brewer.pal(name = "Set3", n = 12))
if(nrow(cit) > length(colors)) {
  cit$color <-  c(colors, colorRampPalette(colors = c("#336666", "#ab3d3f"))(nrow(cit)- length(colors)))  
} else {
  cit$color <- colors[1:nrow(cit)]
}


totF <- df %>% group_by(geo_code) %>% summarise(tot = sum(share)) %>% ungroup() %>% arrange(desc(tot))


wtheme <- function(title.font = "OpenSans-CondensedBold", base.font = "OpenSans-CondensedLight", legend.text.size = 16) {
  theme(
    plot.title = element_text(family = title.font, size = 21, margin = margin(b = 29), colour = "#404040"),
    legend.title = element_text(family = base.font, size = 12, color = "#404040") ,
    legend.text = element_text(family = base.font, size = legend.text.size),
    legend.key.width = unit(1.6, "lines"),
    legend.key.height = unit(1, "lines"),
    legend.position=c(0,1.08), 
    legend.justification=c(0, 0)
  ) 
}

```

## waffle chart

```{r waffle chart}
nrow <- 5
size <- 1
ncol <- floor(max(totF$tot) / nrow)

lang <- 'FR'

for(lang in colnames(txt)) {
  
  # Loop by iso2 and create the 
  wf <- lapply(totF$geo_code, function(iso) {
    cat("\n", iso)
    
    ddd <- filter(df, geo_code == iso, share > 0)
    pc <- round((ddd$FOR[1]/ ddd$TOTAL[1])  * 100, 1)
    # order whether citizen is a EUFTA or not country
    eu.log <- ddd$citizen_code %in% c(eu.ct, "OU") 
    ddd <- ddd[c(which(eu.log), which(!eu.log)),]
    
    # vector: values & translated country names
    dddd <- structure(
      ddd$share, names = countryTranslation(ddd$citizen_code, ifelse(lang == 'AR', 'EN', lang))[,-1])
    # for non-exsistent tranlsation get it from the google sheet
    codes <- unlist(ddd[which(is.na(names(dddd))), "citizen_code"], use.names = F)
    names(dddd)[which(is.na(names(dddd)))] <- txt[codes, lang]
    if(any(is.na(names(dddd)))) {stop ()}
    # add space at the end of the country translation
    names(dddd) <- paste0(names(dddd), "  ")
    
    title <- paste0(ifelse(iso == "EU", txt['EU', lang], countryTranslation(iso, ifelse(lang == 'AR', 'EN', lang))[-1]), "\n")
    pad <- ncol - floor(sum(ddd$share) / nrow)
    
    legend.text.size <- ifelse(lang %in% c('ZH', 'JA'), 12.5, 16)
    
    suppressWarnings(
      waffle(
        dddd, rows = nrow , size=size, pad = pad, 
        colors = cit[match(ddd$citizen_code, cit$citizen_code), 'color'] %>% 
          unlist(use.names = F), legend_pos = "top", 
        title = title, xlab = "\n") +
        wtheme(title.font = txt["title.font", lang], base.font = txt["base.font", lang], legend.text.size = legend.text.size) + 
        guides(fill=guide_legend(nrow=2,byrow=F, title = paste0(pc, txt["subtitle1", lang])))
    )
  })
  
  w.path <- paste0("output/waffle_chart_2016_" , lang, ".png")
  w.html <- paste0("waffle_charts_" , lang, ".html")
  
  png(w.path,  res = 250, pointsize = 12, height = 4700, 
      width = 2000)
  iron( wf[[7]], wf[[1]], wf[[2]], wf[[3]], wf[[4]], wf[[5]], wf[[6]],  wf[[8]], wf[[9]])
  dev.off()
  
  source <- paste0(
    txt['source', lang], ": ", htmlLink(txt['source.url', lang], txt['source.name', lang]), " | ",
    txt['code', lang], ": ", htmlLink(txt['code.url', lang], txt['code.name', lang]), " | ",
    htmlLink("http://www.swissinfo.ch", 'swissinfo.ch'), " | ",
    htmlLink("https://twitter.com/duc_qn", '@duc_qn')    
  )
  footer <- paste0(txt["footer1", lang], "<br>", txt["footer2", lang])
  
  save_html(
    tags$html(
      tags$head(includeHTML("styles.html")),
      tags$body(    
        h2(txt["title1", lang]),
        div(class = "descr", HTML(txt["descr1", lang])),
        h3(txt["h3", lang]),
        div(img(src = basename(w.path), width= "100%")),
        div(class = "footer", HTML(footer)),
        div(id = "cite", HTML(source)),
        HTML(iframeresizer)  
      )), file = w.html
  )
  
  # move stuff
  file.rename(from = w.html, to = paste0("output/", w.html))    
}

    
```

